Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-11-29 dsn-2018/bare_conf.tex
Expanded section on related work
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

From an attacker's perspective, finding a preimage or second preimage has been
difficult. To the author's knowledge, the current bound for a preimage attack
in MD5 is $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further,
while J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather rather long messages
\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
for adversaries, and well past only theoretical attacks. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schlaffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

@@

\subsection{Differential Cryptanalysis}

@@

\subsection{Logical Cryptanalysis}

Logical Cryptanalysis and its encoding as SAT likely started under the work of
F. Massacci in early 2000 with his paper titled ``Logical Cryptanalysis as a
SAT Problem: Encoding and Analysis of the U.S. Data Encryption Standard''.

\subsection{Trustworthy Computing}

@@

>>>>

From an attacker's perspective, finding a preimage or second preimage has been
difficult. To the author's knowledge, the current bound for a preimage attack
in MD5 is $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further,
while J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather long messages
\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
for adversaries, and well past only theoretical attacks. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schlaffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

@@

\subsection{On Differential Cryptanalysis}

\textbf{TODO - expand?}

@@

\subsection{On Logical Cryptanalysis}

\textbf{TODO - expand?}

Logical Cryptanalysis and its encoding as SAT likely started under the work of
F. Massacci in 1999 with his paper titled ``Using walk-SAT and rel-sat for
cryptographic key search''. \cite{Massacci:1999:UWR:1624218.1624261}. However,
much of the early work by F. Massacci was focused on symmetric and asymmetric
ciphers (in \cite{Massacci:1999:UWR:1624218.1624261},
\cite{MASSACCI2000LogicalCA}, and \cite{FIORINI2003101}). It wasn't until the
work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}} in \cite{LogicalAnalysis}
that exploring collisions in the context of SAT was introduced, and until
the work of I. Mironov and L. Zhang in \cite{Mironov2006} that this was
studied for a X. Wang's collision class.

We make use of the \textit{bc2cnf} utility by T. Junttila (available here
\cite{circuits}) as the basis for our models. This allows us to encode hash
functions as generic structures in the circuit description language and
ease the algorithmic creation of models which build on top of them. We then
use CryptoMiniSat5 by M. Soos (\cite{CryptoMiniSat5}) to run the models and
check for a satisfying witness, often an example collision.

With the exception of the work by I. Mironov, the authors find that much of
the previous work in the field has been focused on evaluating the performance
of various SAT solvers. We deviate from this by instead focusing on evaluating
the properties of the hash function--not the SAT solver. We have developed new
techniques for using a SAT solver to inductively build full collisions in MD4
from collisions in reduced-round MD4 and to build addition examples of
collision classes from a single instance of a collision in MD4.

\subsection{On Trustworthy Computing}

@@

\textbf{TODO - introductory paragraph}
